# Building Models {#building-models}

<!-- Models are often similar. -->


In our haste to make progress --- to get all the way through the process of building, interpreting and using models --- we have given short shrift to some of the messy details of model building and evaluation. This chapter fills in those lacunae.


## Transforming predictor variables

It is often convenient to transform a predictor variable. 

### Centering

Recall our model of `income` as a function of `age`. Mathematics:

$$ y_i = \beta_0  + \beta_1 age_i + \epsilon_i$$

We fit this using the `trains` data from **primer.data**.

```{r, message = FALSE}
library(tidyverse)
library(primer.data)
library(rstanarm)
```


```{r}
fit_1 <- stan_glm(formula = income ~ age, 
         data = trains, 
         refresh = 0,
         seed = 9)

print(fit_1, detail = FALSE)
```

There is nothing wrong with this model. Yet the interpretation of $\beta_0$, the intercept in the regression, is awkward. It means the average income for people of age zero. Yet that is useless! There are no people of zero age in our data. And, even if there were, it would be weird to think about such people taking the commuter trade into Boston and filling out our survey forms.

It is easy, however, to *transform* `age` into a variable which makes the intercept more meaningful. Consider a new variable, `c_age`, which is `age` minus the average age in the sample. Using this **c**entered version of age does not change the predictions or residuals in the model, but it does make the intercept easier to interpret.


```{r}
trains_2 <- trains %>% 
  mutate(c_age = age - mean(age))

fit_1_c <- stan_glm(formula = income ~ c_age, 
                    data = trains_2, 
                    refresh = 0,
                    seed = 9)

print(fit_1_c, detail = FALSE)
```

The intercept, `r round(coef(fit_1_c)["(Intercept)"], 0)`, is the expected income for someone with `age_c = 0`, i.e., someone of an average age in the data, which is around `r round(mean(trains$age), 0)`. 

### Scaling

Centering --- changing a vector via addition/subtraction --- often makes the intercept easier to interpret. Scaling --- changing a vector via multiplication/division --- often makes it easier to interpret coefficients. The most common scaling method is to divide the vector by its standard deviation.

```{r}
trains_3 <- trains %>% 
  mutate(s_age = age / sd(age))

fit_1_s <- stan_glm(formula = income ~ s_age, 
                    data = trains_3, 
                    refresh = 0,
                    seed = 9)

print(fit_1_s, detail = FALSE)
```

`s_age` is age **s**caled by its own standard deviation. A change in one unit of `s_age` is the same as a change in one standard deviation of the `age`, which is about `r round(sd(trains$age))`. The interpretation of $\beta_1$ is now:

*When comparing two people, one about 1 standard deviation worth of years older than the other, we expect the older person to earn about `r scales::comma(round(coef(fit_1_s)["age_s"], -3))` more.*

But, because we scaled without centering, the intercept is now back to the (nonsensical) meaning of the expected income for people of age 0. 


### *z*-scores

The most common transformation applies both centering and scaling. The base R function `scale()` subtracts the mean and divides by the standard deviation. A variable so transformed is a "*z*-score, meaning a vector with a mean of zero and a standard deviation of one. Using *z*-scores makes interpretation easier, especially when we seek to compare the importance of different predictors.

```{r}
trains_4 <- trains %>% 
  mutate(z_age = scale(age))

fit_1_z <- stan_glm(formula = income ~ z_age, 
                    data = trains_4, 
                    refresh = 0,
                    seed = 9)

print(fit_1_z, detail = FALSE)
```

The two parameters are easy to interpret after this transformation.

*The expected income of someone of average age, which is about `r round(mean(trains$age))` in this study, is about `r round(coef(fit_1_z)["(Intercept)"], 0)` dollars.*

*When comparing two individuals who differ in age by one standard deviation, which is about `r round(sd(trains$age))` in this study, the older person is expected to earn about `r `scales::comma(round(coef(fit_1_z)["age_z"], -3))` more than the younger.*

Note that, when using *z*-scores, we would often phrase this comparison in terms of "sigmas." One person is "one sigma" older than another person means that they are one standard deviation older. This is simple enough, once you get used to it, but also confusing since we already using the word "sigma" to mean $\sigma$, the standard deviation of $\epsilon_i$. Alas, *language is something we deal with rather than control.* You will hear the same word "sigma" applied to both concepts, even in the same sentence. Determine meaning by context.

### Taking logs

<!-- * taking logs. When? Why? Doing so on dependent variable is a much bigger deal than doing it on independent variables since the former changes he very essence of the model. You can't really (?) compare two different models if one has taken logs of Y and one has not. -->

### Adding transformed terms

Instead of simply transforming variables, we can add more terms which are transformed versions of a variable. Consider the relation of `height` to `age` in `nhanes`. Let's start by dropping the missing values.

```{r}
no_na_nhanes <- nhanes %>% 
  select(height, age) %>% 
  drop_na() 
```

Fit and plot a simple linear model:

```{r}
nhanes_1 <- stan_glm(height ~ age,
                     data = no_na_nhanes,
                     refresh = 0,
                     seed = 47)

no_na_nhanes %>% 
  ggplot(aes(x = age, y = height)) +
    geom_point(alpha = 0.1) +
    geom_line(aes(y = predict(nhanes_1)), 
              color = "red") +
    labs(title = "Age and Height",
         subtitle = "Children are shorter, but a linear fit is poor",
         x = "Age",
         y = "Height (cm)",
         caption = "Data source: NHANES")
```

That is not a very good model, obviously.

Adding a quadratic term makes it better. (Note the need for `I()` in creating the squared term within the `formula` argument.)


```{r}
nhanes_2 <- stan_glm(height ~ age + I(age^2),
                     data = no_na_nhanes,
                     refresh = 0,
                     seed = 33)

no_na_nhanes %>% 
  ggplot(aes(x = age, y = height)) +
    geom_point(alpha = 0.1) +
    geom_line(aes(y = predict(nhanes_2)), 
              color = "red") +
    labs(title = "Age and Height",
         subtitle = "Quadratic fit is much better, but still poor",
         x = "Age",
         y = "Height (cm)",
         caption = "Data source: NHANES")
```

Still, we have not made use of our background knowledge in creating these variables. We know that people don't get any taller after age 18 or so. Let's create variables which capture that break.

```{r}
nhanes_3 <- stan_glm(height ~ I(ifelse(age > 18, 18, age)),
                     data = no_na_nhanes,
                     refresh = 0,
                     seed = 23)

no_na_nhanes %>% 
  ggplot(aes(x = age, y = height)) +
    geom_point(alpha = 0.1) +
    geom_line(aes(y = predict(nhanes_2)), 
              color = "red") +
    labs(title = "Age and Height",
         subtitle = "Domain knowledge makes for better models",
         x = "Age",
         y = "Height (cm)",
         caption = "Data source: NHANES")
```

The point is that we should not take the variables we receive as *given*. We are the captains of our souls. We transform variables as needed.



### If parameters are unicorns, why bother with transformations?

<!-- * standardizing variables. Key advantage is to make the coefficients easier to interpret. Even though the "right" thing to do in looking at models is to use posterior_* to look at model predictions, this takes some time. It is very handy, first pass, to look at all the coefficients and get a sense of which ones are "interesting." Then, focus on those when making newobs tibbles. -->

## Transforming the outcome variable

Transforming predictor variables is generally uncontroversial. It does not matter much. Change most continuous predictor variables to $z$-scores and you won't go far wrong. Or keep them in their original form, and take care with your interpretations. It's all good.

Transforming the outcome variable is a much more difficult question, to which there are few right answers.




## Selecting variables


How do we decide which variables to include in a model? There is not

<!-- * Keeping or throwing away variables. How do we decide? If something is big and significant, we always keep it. If it is small and insignificant, we don't, unless our audience really wants to see it. If a mix, reasonable people differ. No right answer! -->

<!-- * Creating new variables, especially by combining different ones into one combined variable. -->


## Comparing models

Deciding which variables to include in a model is a subset of the larger question: How do we decide which model, out of the set of possible models, to choose?

Consider two models which explain attitudes to immigration among Boston commuters.

```{r}
fit_liberal <- stan_glm(formula = att_end ~ liberal,
                  data = trains,
                  refresh = 0,
                  seed = 42)

print(fit_liberal, detail = FALSE)
```


```{r}
fit_att_start <- stan_glm(formula = att_end ~ att_start,
                  data = trains,
                  refresh = 0,
                  seed = 85)

print(fit_att_start, detail = FALSE)
```

They both seem like good models! The results make sense. People who are liberal have more liberal attitudes about immigration, so we would expect their `att_end` scores to be lower. We would also expect people to provide similar answers in two surveys administered a week or two apart. It makes sense that those with higher (more conservative) values for `att_start` would also have higher values for `att_end`.

How do we choose between these models?

### Better models make better predictions

The most obvious criteria for comparing models is the accuracy of the predictions. For example, consider the use of `liberal` to predict `att_end`.


```{r}
trains %>% 
  mutate(pred_liberal = predict(fit_liberal)) %>% 
  ggplot(aes(x = pred_liberal, y = att_end)) +
    geom_jitter(width = 0.05, height = 0.2, alpha = 0.5) +
    labs(title = "Modeling Attitude Toward Immigration",
         subtitle = "Liberals are less conservative",
         x = "Predicted Attitude",
         y = "True Attitude")
```

<!-- DK: Could put this discussion in fig.cap. Could add some details to the plots to indicate the points that I am talking about. Should create an animation, which brings in new point, forecasts it, and then measures how wrong the model was. -->

Because there are only two possible values for `liberal` --- TRUE and FALSE --- there are only two predictions which this model will make: about 10 for `liberal == FALSE` and about 8 for `liberal == TRUE`. (The points in the above plot are jittered.)  For some individuals, these are perfect predictions. For others, they are poor predictions. (Note the two individuals who are `liberal == TRUE`, and who the model thinks will have `att_end == 8`, but who have `att_end == 15`. The mode got them both very, very wrong.)

Consider our second model, using `att_start` to forecast `att_end`.

```{r}
trains %>% 
  mutate(pred_liberal = predict(fit_att_start)) %>% 
  ggplot(aes(x = pred_liberal, y = att_end)) +
    geom_jitter(width = 0.05, height = 0.2, alpha = 0.5) +
    labs(title = "Modeling Attitude Toward Immigration",
         subtitle = "Survey responses are somewhat consistent",
         x = "Predicted Attitude",
         y = "True Attitude")
```


Because `att_end` takes on `r length(unique(trains$att_end))` unique values, the model makes `r length(unique(trains$att_end))` unique predictions. Some of those predictions are perfect! But others are very wrong. Note the individual with a predicted `att_end` of around 9 but with an actual value of `15`. That is a big miss!


Rather than looking at individual cases, we need to look at the errors for all the predictions. Fortunately, a prediction error is the same thing as a residual, which is easy enough to calculate.

```{r}
trains %>% 
  select(att_end, att_start, liberal) %>% 
  mutate(pred_lib = predict(fit_liberal)) %>% 
  mutate(resid_lib = predict(fit_liberal) - att_end) %>% 
  mutate(pred_as = predict(fit_att_start)) %>% 
  mutate(resid_as = predict(fit_att_start) - att_end)
```

Let's look at the square root of the average squared error. 

```{r}
trains %>% 
  select(att_end, att_start, liberal) %>% 
  mutate(lib_err = (predict(fit_liberal) - att_end)^2) %>% 
  mutate(as_err = (predict(fit_att_start) - att_end)^2) %>% 
  summarize(lib_sigma = sqrt(mean(lib_err)),
            as_sigma = sqrt(mean(as_err))) 
```

There are many different measures of the error which we might calculate. The squared difference is most common for historical reasons: it was the mathemtically most tractable in the pre-computer age. Having calculated a squared difference for each observation, we can sum them or take their average or take the square root of their average. All produce the same relative ranking, but the last is most popular because it (more or less) corresponds to the estimated $\sigma$ for a linear model. Note how these measures are the same as the ones produced by the Bayesian models created above.

<!-- DK: Could make a movie where we get an X value, we calculate a prediction, we plot it, then we plot the truth, then we measure the distance, then we calculate a squared error. Then speed up the movie, doing it for all the points.  -->

<!-- DK: Could discuss uncertainty in the estimated sigma. But left it out in this draft. -->

Sadly, it is not wise to simply select the model which fits the data best because doing so can be misleading. After all, you are cheating! You are using that very data to select your parameters and then, after using the data once, turning around and "checking" to see how well your model fits the data. It better fit! You used it to pick your parameters! This is the danger of *overfitting*.


### Beware overfitting

One of the biggest dangers in data science is *overfitting*, using a model with too many parameters which fits the data we have too well and, therefore, works poorly on data we have yet to see. Consider a simple example with 10 data points.

```{r, echo = FALSE}
ovrftng <- tibble(x = 1:10,
                  y = c(1.9, 1.9, 4.9, 2.7, 2.6, 
                        5.2, 6.8, 8.9, 9.0, 9.9))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    labs(title = "Create a Model which Predicts `y` Given `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

What happens when we fit a model with one predictor?

```{r, echo = FALSE}
one_pred <- lm(y ~ x,
               data = ovrftng)
  
newdata <- tibble(x = seq(1, 10, by = 0.01),
                  y = predict(one_pred, 
                              newdata = tibble(x = x)))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    geom_line(data = newdata, 
              aes(x, y)) +
    labs(title = "`y` as a Linear Function of `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

That is a reasonable model. It does not fit the data particularly well, but we certainly believe that higher values of `x` are associated with higher values of `y`. A linear fit is not unreasonable.

But we can also use some of the lessons from above and try a quadratic fit by adding $x^2$ as a predictor.

```{r, echo = FALSE}
two_pred <- lm(y ~ poly(x, 2),
               data = ovrftng)

newdata <- tibble(x = seq(1, 10, by = 0.01),
                  y = predict(two_pred, 
                              newdata = tibble(x = x)))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    geom_line(data = newdata, 
              aes(x, y)) +
    labs(title = "`y` as a Quadratic Function of `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

Is this a better model? Maybe? 

But why stop at adding $x^2$ to the regression? Why not add $x^3$, $x^4$ and all the way to $x^9$? When we do so, the fit is much better.


```{r}
nine_pred <- lm(y ~ poly(x, 9),
                       data = ovrftng)

newdata <- tibble(x = seq(1, 10, by = 0.01),
                  y = predict(nine_pred, 
                              newdata = tibble(x = x)))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    geom_line(data = newdata, 
              aes(x, y)) +
    labs(title = "`y` as a 9-Degree Polynomial Function of `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

*If the only criteria we cared about was how well the model predicts using the data on which the parameters were estimated, then a model with more parameters will always be better.* But that is not what truly matters. What matters is how well the model works on data which was not used to create the model.


### Better models make better predictions on new data



### Selecting the rows

## Testing is nonsense


## Summary


