# Four Parameters {#four-parameters}

<!-- Explore causal effects of models following chapter 18/19 in ROAS. Regress ending att on starting attitude and treatment effect. Show parallel lines. That is four parameters! Can also do the same thing with income and treatment. That will be a much worse model. Save interaction for five parameter chapter. -->

<!-- Add page 217 to selecting variables. Simple version. Can't talk about everything. Can't talk about interactions, for example. Talk about testing and how stupid it is! MF: Done.-->

<!-- Write out better models make better predictions on new data. Concepts! Cross validation. Leave one out. Blah, blah, blah. MF: Done. -->


<!-- In Better models make better predictions, add red dots (to first) and red line (to second) to indicate where all the data points would be if your model were perfect. Should we change the aspect on these plots so that they are "square"? Maybe add y and y_hat to the x-axis labels for these two plots. At least discuss in the text. These plots are weird! Not seen before! MF: Done. -->


*This chapter is still a DRAFT. Check back in a few weeks.*


In our haste to make progress --- to get all the way through the process of building, interpreting and using models --- we have given short shrift to some of the messy details of model building and evaluation. This chapter fills in those lacunae. We will also introduce models with four parameters. 


## Transforming variables

<!-- Look at "Scaling of predictors and regression coefficients" in RAOS (pg. 183) 

Do we discuss anything in the "Other transformations section" (pg. 195)? This includes square root transformations, idiosyncratic transformations, & using discrete vs continuous predictors. DK: Not square root. Other stuff? Maybe at the end? -->

An important question for us data scientists: what happens when our assumptions break down? The first option is to extend our model, perhaps by addressing our concerns with validity by including error models. The second option, which we will focus on here, is to modify our data or model so that our assumptions are more reasonable. How do we accomplish this?

It is often convenient to transform a predictor variable so that our model makes more sense. Let's explore exactly what that means. 

### Centering

Recall our model of `income` as a function of `age`. Mathematics:

$$ y_i = \beta_0  + \beta_1 age_i + \epsilon_i$$

We fit this using the `trains` data from **primer.data**.

```{r, message = FALSE}
library(tidyverse)
library(primer.data)
library(rstanarm)
```


```{r}
fit_1 <- stan_glm(formula = income ~ age, 
         data = trains, 
         refresh = 0,
         seed = 9)

print(fit_1, detail = FALSE)
```

There is nothing wrong with this model. Yet the interpretation of $\beta_0$, the intercept in the regression, is awkward. It represents the average income for people of age zero. That is useless! There are no people of zero age in our data. And, even if there were, it would be weird to think about such people taking the commuter train into Boston and filling out our survey forms.

It is easy, however, to *transform* `age` into a variable which makes the intercept more meaningful. Consider a new variable, `c_age`, which is `age` minus the average age in the sample. Using this **c**entered version of age does not change the predictions or residuals in the model, but it does make the intercept easier to interpret.


```{r}
trains_2 <- trains %>% 
  mutate(c_age = age - mean(age))

fit_1_c <- stan_glm(formula = income ~ c_age, 
                    data = trains_2, 
                    refresh = 0,
                    seed = 9)

print(fit_1_c, detail = FALSE)
```

The intercept, `r scales::comma(round(coef(fit_1_c)["(Intercept)"], 0))`, is the expected income for someone with `age_c = 0`, i.e., someone of an average age in the data, which is around `r round(mean(trains$age), 0)`. 

### Scaling

Centering --- changing a variable via addition/subtraction --- often makes the intercept easier to interpret. Scaling --- changing a variable via multiplication/division --- often makes it easier to interpret coefficients. The most common scaling method is to divide the variable by its standard deviation.

```{r}
trains_3 <- trains %>% 
  mutate(s_age = age / sd(age))

fit_1_s <- stan_glm(formula = income ~ s_age, 
                    data = trains_3, 
                    refresh = 0,
                    seed = 9)

print(fit_1_s, detail = FALSE)
```

`s_age` is age **s**caled by its own standard deviation. A change in one unit of `s_age` is the same as a change in one standard deviation of the `age`, which is about `r round(sd(trains$age))`. The interpretation of $\beta_1$ is now:

*When comparing two people, one about 1 standard deviation worth of years older than the other, we expect the older person to earn about `r scales::comma(round(coef(fit_1_s)["s_age"], -3))` more.*

But, because we scaled without centering, the intercept is now back to the (nonsensical) meaning of the expected income for people of age 0. 


### *z*-scores

The most common transformation applies both centering and scaling. The base R function `scale()` subtracts the mean and divides by the standard deviation. A variable so transformed is a "*z*-score," meaning a variable with a mean of zero and a standard deviation of one. Using *z*-scores makes interpretation easier, especially when we seek to compare the importance of different predictors.

```{r}
trains_4 <- trains %>% 
  mutate(z_age = scale(age))

fit_1_z <- stan_glm(formula = income ~ z_age, 
                    data = trains_4, 
                    refresh = 0,
                    seed = 9)

print(fit_1_z, detail = FALSE)
```

The two parameters are easy to interpret after this transformation.

*The expected income of someone of average age, which is about `r round(mean(trains$age))` in this study, is about `r scales::comma(round(coef(fit_1_z)["(Intercept)"], -3))` dollars.*

*When comparing two individuals who differ in age by one standard deviation, which is about `r round(sd(trains$age))` years in this study, the older person is expected to earn about `r scales::comma(round(coef(fit_1_z)["z_age"], -3))` more than the younger.*

Note that, when using *z*-scores, we would often phrase this comparison in terms of "sigmas." One person is "one sigma" older than another person means that they are one standard deviation older. This is simple enough, once you get used to it, but also confusing since we are already using the word "sigma" to mean $\sigma$, the standard deviation of $\epsilon_i$. Alas, *language is something we deal with rather than control.* You will hear the same word "sigma" applied to both concepts, even in the same sentence. Determine meaning by context.

### Taking logs

It is often helpful to take the log of predictor variables, especially in cases in which their distribution is skewed. You should generally only take the log of variables for which all the values are strictly positive. The log of a negative number is not defined. Consider the number of registered voters (`rv13`) at each of the polling stations in `kenya`. 

```{r}
x <- kenya %>% 
  filter(rv13 > 0)

rv_p <- x %>% 
  ggplot(aes(rv13)) + 
    geom_histogram(bins = 100) +
    labs(x = "Registered Voters",
         y = NULL) 

log_rv_p <- x %>% 
  ggplot(aes(log(rv13))) + 
    geom_histogram(bins = 100) +
    labs(x = "Log of Registered Voters",
         y = NULL) +
    expand_limits(y = c(0, 175))

rv_p + log_rv_p +
  plot_annotation(title = 'Registered Votes In Kenya Communities',
                  subtitle = "Taking logs helps us deal with outliers")
```

Most experienced data scientists would use the log of `rv13` rather than the raw value. Comments:

* We do not know the "true" model. Who is to say that a model using the raw value is right or wrong?

* Check whether or not this choice meaningfully affects the answer to your question. Much of the time, it won't. That is, our inferences are often fairly "robust" to small changes in the model. If you get the same answer with `rv13` as from `log_rv13`, then no one cares which you use.

* Follow the conventions in your field. If everyone does X, then you should probably do X, unless you have a good reason not to. If you do have such a reason, explain it prominently.

* Most professionals, when presented with data distributed like `rv13`, would take the log. Professionals hate (irrationally?) outliers. Any transformation which makes a distribution look more normal is generally considered a good idea.

Many of these suggestions apply to every aspect of the modeling process.

<!-- Split out general advice to elsewhere. -->

### Adding transformed terms

Instead of simply transforming variables, we can add more terms which are transformed versions of a variable. Consider the relation of `height` to `age` in `nhanes`. Let's start by dropping the missing values.

```{r}
no_na_nhanes <- nhanes %>% 
  select(height, age) %>% 
  drop_na() 
```

Fit and plot a simple linear model:

```{r}
nhanes_1 <- stan_glm(height ~ age,
                     data = no_na_nhanes,
                     refresh = 0,
                     seed = 47)

no_na_nhanes %>% 
  ggplot(aes(x = age, y = height)) +
    geom_point(alpha = 0.1) +
    geom_line(aes(y = fitted(nhanes_1)), 
              color = "red") +
    labs(title = "Age and Height",
         subtitle = "Children are shorter, but a linear fit is poor",
         x = "Age",
         y = "Height (cm)",
         caption = "Data source: NHANES")
```

That is not a very good model, obviously.

Adding a quadratic term makes it better. (Note the need for `I()` in creating the squared term within the `formula` argument.)


```{r}
nhanes_2 <- stan_glm(height ~ age + I(age^2),
                     data = no_na_nhanes,
                     refresh = 0,
                     seed = 33)

no_na_nhanes %>% 
  ggplot(aes(x = age, y = height)) +
    geom_point(alpha = 0.1) +
    geom_line(aes(y = fitted(nhanes_2)), 
              color = "red") +
    labs(title = "Age and Height",
         subtitle = "Quadratic fit is much better, but still poor",
         x = "Age",
         y = "Height (cm)",
         caption = "Data source: NHANES")
```

Still, we have not made use of our background knowledge in creating these variables. We know that people don't get any taller after age 18 or so. Let's create variables which capture that break.

```{r}
nhanes_3 <- stan_glm(height ~ I(ifelse(age > 18, 18, age)),
                     data = no_na_nhanes,
                     refresh = 0,
                     seed = 23)

no_na_nhanes %>% 
  ggplot(aes(x = age, y = height)) +
    geom_point(alpha = 0.1) +
    geom_line(aes(y = fitted(nhanes_3)), 
              color = "red") +
    labs(title = "Age and Height",
         subtitle = "Domain knowledge makes for better models",
         x = "Age",
         y = "Height (cm)",
         caption = "Data source: NHANES")
```

The point is that we should not take the variables we receive as *given*. We are the captains of our souls. We transform variables as needed.



<!-- ### If parameters are unicorns, why bother with transformations? -->

<!-- * standardizing variables. Key advantage is to make the coefficients easier to interpret. Even though the "right" thing to do in looking at models is to use posterior_* to look at model predictions, this takes some time. It is very handy, first pass, to look at all the coefficients and get a sense of which ones are "interesting." Then, focus on those when making newobs tibbles. -->

### Transforming the outcome variable

Transforming predictor variables is uncontroversial. It does not matter much. Change most continuous predictor variables to $z$-scores and you won't go far wrong. Or keep them in their original form, and take care with your interpretations. It's all good.

Transforming the outcome variable is a much more difficult question. Imagine that we seek to create a model which explains `rv13` from the `kenya` tibble. Should we transform it?

* Maybe? There are no right answers. A model with `rv13` as the outcome variable is different from a model with `log(rv13)` as the outcome. The two are not directly comparable.

* Much of the same advice with regard to taking logs of predictor variables applies here as well.


See @roas


<!-- * Creating new variables, especially by combining different ones into one combined variable. -->

## Selecting variables

How do we decide which variables to include in a model? There is no one right answer to this question. 

### General guidelines for selecting variables 

When deciding which variables to keep or discard in our models, our advice is to keep a variable X if any of the following circumstances apply:

* The variable has a large and well-estimated coefficient. This means, roughly, that the 95% confidence interval excludes zero.

* Underlying theory/observation suggests that X has a meaningfully connection to the outcome variable.

* If the variable has a small standard error, it is general practice to include it in our model to improve predictions. If the standard error of a coefficient is large and we find no other reason to include it in our model, it may be a signal to remove the variable from your model. 

* If a predictor isn't statistically significant, we would normally consider discarding it. The exception to this rule is if we wanted to compare two indicators (for instance, if we wanted to compare results from the variable `party` -- to look at Democrats versus Republicans -- in the `trains` dataset) in spite of that variable not appearing to be significant. 

* It is standard in your field to include X in such regressions.

* Your boss/client/reviewer/supervisor wants to include X.

Another trick to consider is the Divide-by-Four rule: As a rule of convenience, we can take logistic regression coefficients (other than the constant term) and divide them by 4 to get an upper bound of the predictive difference corresponding to a unit difference in x. All this means is that, when evaluating if a predictor is helpful, divide the coefficient by four. If the predictor exceeds this value, it is likely to be significant. 

Let's use the `trains` dataset to evaluate how helpful certain variables are to creating an effective model, based on the guidelines above.

### Variables in the `trains` dataset

To look at our recommendations in practice, let's focus on the `trains` dataset. The variables in `trains` include `gender`, `liberal`, `party`, `age`, `income`, `att_start`, `treatment`, and `att_end`. Which variables would be best to include in a model? 

### att_end ~ treatment + att_start

First, let's look at a model with a left hand variable, `att_end`, and two right side variables, `treatment` and `att_start`.

```{r}
fit_1 <- stan_glm(att_end ~ treatment + att_start, 
                    data = trains, 
                    refresh = 0)

fit_1
```

How do we decide which variables are useful? First, let's look at our intercept. The 95% confidence interval for  `att_end` is equal to the coefficient plus or minus 2 standard errors = 2.3 +/- 0.8 = 1.5 to 3.1. When we divide this by four, we get .375 to .775. Therefore, any variable that exceeds these values is likely to be significant enough to include in our model. 

As `treatmentControl`, a value which impacts our (Intercept) by a factor of -1.0, not only exceeds the minimum requirement of .375, but also exceeds the upper boundary of significance of .775, we definitely want to keep the variable `treatment`. In addition to being significant, it satisfies a number of other requirements:
- The variable has a small standard error.
- The variable is considered an indicator variable, which separates two groups of significance (treatment and control) that we would like to study. 

The variable `att_start`, with a coefficient of 0.8, is also significant. In addition, the standard error is small, the impact on our prediction for `att_end` is meaningful, and it is standard practice to include a "before" comparison to our "after" values. 

**Conclusion**: keep both variables! `Treatment` and `att_start` are both significant, as well as satisfying other requirements in our guidelines. They are *more* than worthy of inclusion in our model.  

### income ~ age + liberal

Now, we will look at `income` as a function of `age` and `liberal`, a proxy for political party. 

```{r}
fit_2 <- stan_glm(income ~ age + liberal, 
                    data = trains, 
                    refresh = 0)

fit_2
```

Great! We have an estimate for income of those who fall into the category of `liberalFALSE`, as well as data on our right hand side variables of `age` and `liberalTRUE`. Recall the divide-by-four rule. The range for (Intercept), our 95% confidence interval, is calculated by taking our coefficient plus or minus 2 standard errors (MAD_SD). The lower range is $58703.7$ and the upper range is $162834.9$. Dividing both of these values by 4, we get $14675.93$ and $40708.725$. 

As we can see, `liberalTRUE` falls within the boundaries for significance. It is, therefore, a great addition to our model! 

The variable `age`, however, does not seem to be significant. The value of `age` is very low. It does not seem to make a meaningful difference. This being said, we just learned about a bunch of helpful transformations! We also know that the variable `age`, unmodified, is calculating starting at age 0, an impossibility in our data. 

Recall the work above that z-scaled the variable `age` to be more significant. This is pasted below for reference. 

```{r}
trains_age <- trains %>% 
  mutate(z_age = scale(age))
```

Let's run our regression again, this time using `z_age`, to see if the resulting value is more significant. 

```{r}
fit_scaled <- stan_glm(income ~ z_age + liberal, 
                    data = trains_age, 
                    refresh = 0)

fit_scaled
```

This looks better! Even with our transformation, `z_age` is still on the lower bounds of our significance. Whether we include it in our model is not clear, especially when compared with the variable `liberal`, which shows clear significance. 

**Conclusion**: definitely keep `liberal`! Age is less clear. Age when z-scaled is *just* significant enough to consider including. It is really a matter of preference at this point. If age is important to our model in some way, we keep it. If it isn't relevant to our conclusions, we consider discarding it. 

### liberal ~ I(income/1e+05) + party

We will now look at liberal as a function of a transformed `income` and `party`. The reason we have transformed `income` here is due to the fact that, without a transformation, income makes predictions according to *very small income disparities*. This is not helpful. To enhance the significance of income, we have divided the variable by $100000$. (Note the need for `I()` in creating the income term within the `formula` argument.)


```{r}
fit_3 <- stan_glm(liberal ~ I(income/1e+05) + party, 
                    data = trains, 
                    refresh = 0)

fit_3
```

Let's start by looking at our (Intercept). The 95% confidence interval is $0.7$ plus or minus 2 standard errors ($0.2$). Therefore, the confidence interval is $0.5$ to $0.9$. Divided by four, we get $0.125$ to $0.225$.

As we can see instantly, `partyRepublican` is very meaningful. We will want this variable in our model. 

The variable `income`, however, is still not very significant. The `MAD_SD` for income is also equal to the median itself, suggesting a large standard error. Though we may choose to include `income` for reasons unrelated to significance, it does not appear to be worthy of inclusion on the basis of significance alone. 

**Conclusion**: the variable `party` is significant; we will include it in our model. Unless we have a strong reason to include `income`, it may not be helpful in creating our models. 

### Final thoughts

Now that we have looked at three cases of variables and decided whether they should be included, let's discuss the concept of selecting variables generally. 

The variables we decided to keep and discard are *not necessarily the variables you would keep or discard*, or the variables that any other data scientist would keep or discard. It is much easier to **keep** a variable than it is to build a case for discarding a variable. Variables are helpful even when not significant, particularly when they are indicator variables which may separate the data into two groups that we want to study. 

The process of selecting variables -- though we have guidelines -- is complicated. There are many reasons to include a variable in a model. The main reasons to exclude a variable are if the variable isn't significant or if the variable has a large standard error. 

<!-- Introduce 4 parameter models. Use the exact same models as in chapter 8, but add a second right hand side variable. At the start, only use one categorical and one numeric. MF: Done.-->

<!-- Discuss indicator & index variables prior to explaining guidelines? Not this week! -->


<!-- Go through one of the datasets and explain whether we would include variables X, Y, and Z. Not this week.  Try out several different left-hand side variables, like age and att_end, which were teh dependent variables that we used in Chapter 8. In this chapter, we will allow for two right-hand side variables. (Perhaps these should always be one numeric and one categorical, except for the last one.) Connect to examples from chapter 8. Introduce four parameter models. MF: Done. --> 

<!-- First case att_end ~ treatment + att_start. Keep both! Second case income ~ age + liberal. Strong keep for liberal. age is trickier. It is small. (Not really! Depends on whether or not you are looking at one year or 10. Don't let arbitrary scalings confuse you. Measured in decades it is 10 times bigger. Discuss z scaling?)  Not quite significant. No right answer! Third and last case is logistic, liberal ~ I(income/1e+05) + party. Tricky! Maybe keep both? But stronger case for party. Maybe discuss interpretation of coefficients of logistic regressions? Divide by 4 rule? MF: Done.-->


## Comparing models in theory

Deciding which variables to include in a model is a subset of the larger question: How do we decide which model, out of the set of possible models, to choose?

Consider two models which explain attitudes to immigration among Boston commuters.

```{r}
fit_liberal <- stan_glm(formula = att_end ~ liberal,
                  data = trains,
                  refresh = 0,
                  seed = 42)

print(fit_liberal, detail = FALSE)
```


```{r}
fit_att_start <- stan_glm(formula = att_end ~ att_start,
                  data = trains,
                  refresh = 0,
                  seed = 85)

print(fit_att_start, detail = FALSE)
```

They both seem like good models! The results make sense. People who are liberal have more liberal attitudes about immigration, so we would expect their `att_end` scores to be lower. We would also expect people to provide similar answers in two surveys administered a week or two apart. It makes sense that those with higher (more conservative) values for `att_start` would also have higher values for `att_end`.

How do we choose between these models?

### Better models make better predictions

The most obvious criteria for comparing models is the accuracy of the predictions. For example, consider the use of `liberal` to predict `att_end`.


```{r}
trains %>% 
  mutate(pred_liberal = fitted(fit_liberal)) %>% 
  ggplot(aes(x = pred_liberal, y = att_end)) +
    geom_jitter(width = 0.05, height = 0.2, alpha = 0.5) +
  
  # Add a red circle where our predictions are most accurate (where the x and y
  # values are the same, which is where our predictions = the true attitudes).
  # pch = 1 makes the inside of the point translucent to show the number of
  # correct predictions.
  
  geom_point(aes(x = 8, y = 8), 
             size = 20, pch = 1, 
             color = "red") +
  geom_point(aes(x = 10, y = 10), 
             size = 20, pch = 1, 
             color = "red") +
    labs(title = "Modeling Attitude Toward Immigration",
         subtitle = "Liberals are less conservative",
         x = "Predicted Attitude",
         y = "True Attitude")
```

<!-- DK: Could put this discussion in fig.cap. Could add some details to the plots to indicate the points that I am talking about. Should create an animation, which brings in new point, forecasts it, and then measures how wrong the model was. -->

Because there are only two possible values for `liberal` --- TRUE and FALSE --- there are only two predictions which this model will make: about 10 for `liberal == FALSE` and about 8 for `liberal == TRUE`. (The points in the above plot are jittered.)  For some individuals, these are perfect predictions. For others, they are poor predictions. the red circles on our plot illustrate the areas where our predictions are equal to true values. As we can see, the model isn't great at predicting attitude end. (Note the two individuals who are `liberal == TRUE`, and who the model thinks will have `att_end == 8`, but who have `att_end == 15`. The model got them both very, very wrong.)

Consider our second model, using `att_start` to forecast `att_end`.

```{r}
trains %>% 
  mutate(pred_liberal = fitted(fit_att_start)) %>% 
  ggplot(aes(x = pred_liberal, y = att_end)) +
    geom_jitter(width = 0.05, height = 0.2, alpha = 0.5) +
  
  # Insert red line where our predictions = the truth using geom_abline with an
  # intercept, slope, and color.
  
  geom_abline(intercept = 0, slope = 1, color = "red") +
    labs(title = "Modeling Attitude Toward Immigration",
         subtitle = "Survey responses are somewhat consistent",
         x = "Predicted Attitude",
         y = "True Attitude")
```


Because `att_end` takes on `r length(unique(trains$att_end))` unique values, the model makes `r length(unique(trains$att_end))` unique predictions. Some of those predictions are perfect! But others are very wrong. The red line shows the points where our predictions match the truth. Note the individual with a predicted `att_end` of around 9 but with an actual value of `15`. That is a big miss!

Rather than looking at individual cases, we need to look at the errors for all the predictions. Fortunately, a prediction error is the same thing as a residual, which is easy enough to calculate.

```{r}
trains %>% 
  select(att_end, att_start, liberal) %>% 
  mutate(pred_lib = fitted(fit_liberal)) %>% 
  mutate(resid_lib = fitted(fit_liberal) - att_end) %>% 
  mutate(pred_as = fitted(fit_att_start)) %>% 
  mutate(resid_as = fitted(fit_att_start) - att_end)
```

Let's look at the square root of the average squared error. 

```{r}
trains %>% 
  select(att_end, att_start, liberal) %>% 
  mutate(lib_err = (fitted(fit_liberal) - att_end)^2) %>% 
  mutate(as_err = (fitted(fit_att_start) - att_end)^2) %>% 
  summarize(lib_sigma = sqrt(mean(lib_err)),
            as_sigma = sqrt(mean(as_err))) 
```

There are many different measures of the error which we might calculate. The squared difference is most common for historical reasons: it was the mathematically most tractable in the pre-computer age. Having calculated a squared difference for each observation, we can sum them or take their average or take the square root of their average. All produce the same relative ranking, but the last is most popular because it (more or less) corresponds to the estimated $\sigma$ for a linear model. Note how these measures are the same as the ones produced by the Bayesian models created above.

<!-- DK: Could make a movie where we get an X value, we calculate a prediction, we plot it, then we plot the truth, then we measure the distance, then we calculate a squared error. Then speed up the movie, doing it for all the points.  -->

<!-- DK: Could discuss uncertainty in the estimated sigma. But left it out in this draft. -->

Sadly, it is not wise to simply select the model which fits the data best because doing so can be misleading. After all, you are cheating! You are using that very data to select your parameters and then, after using the data once, turning around and "checking" to see how well your model fits the data. It better fit! You used it to pick your parameters! This is the danger of *overfitting*.


### Beware overfitting

One of the biggest dangers in data science is *overfitting*, using a model with too many parameters which fits the data we have too well and, therefore, works poorly on data we have yet to see. Consider a simple example with 10 data points.

```{r, echo = FALSE}
ovrftng <- tibble(x = 1:10,
                  y = c(1.9, 1.9, 4.9, 2.7, 2.6, 
                        5.2, 6.8, 8.9, 9.0, 9.9))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    labs(title = "Create a Model which Predicts `y` Given `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

What happens when we fit a model with one predictor?

```{r, echo = FALSE}
one_pred <- lm(y ~ x,
               data = ovrftng)
  
newdata <- tibble(x = seq(1, 10, by = 0.01),
                  y = predict(one_pred, 
                              newdata = tibble(x = x)))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    geom_line(data = newdata, 
              aes(x, y)) +
    labs(title = "`y` as a Linear Function of `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

That is a reasonable model. It does not fit the data particularly well, but we certainly believe that higher values of `x` are associated with higher values of `y`. A linear fit is not unreasonable.

<!-- DK: Add a discussion about the difference between plotting y versus x, as we do here, and y versus y_hat, which is what we did in the previous section. In fact, show y versus y_hat with this model, right now! -->

But we can also use some of the lessons from above and try a quadratic fit by adding $x^2$ as a predictor.

```{r, echo = FALSE}
two_pred <- lm(y ~ poly(x, 2),
               data = ovrftng)

newdata <- tibble(x = seq(1, 10, by = 0.01),
                  y = predict(two_pred, 
                              newdata = tibble(x = x)))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    geom_line(data = newdata, 
              aes(x, y)) +
    labs(title = "`y` as a Quadratic Function of `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

Is this a better model? Maybe? 

But why stop at adding $x^2$ to the regression? Why not add $x^3$, $x^4$ and all the way to $x^9$? When we do so, the fit is much better.


```{r}
nine_pred <- lm(y ~ poly(x, 9),
                       data = ovrftng)

newdata <- tibble(x = seq(1, 10, by = 0.01),
                  y = predict(nine_pred, 
                              newdata = tibble(x = x)))

ovrftng %>% 
  ggplot(aes(x, y)) +
    geom_point() +
    geom_line(data = newdata, 
              aes(x, y)) +
    labs(title = "`y` as a 9-Degree Polynomial Function of `x`") +
    scale_x_continuous(breaks = seq(2, 10, 2)) +
    scale_y_continuous(breaks = seq(2, 10, 2)) 
```

*If the only criteria we cared about was how well the model predicts using the data on which the parameters were estimated, then a model with more parameters will always be better.* But that is not what truly matters. What matters is how well the model works on data which was not used to create the model.


<!-- DK: Discuss how this is the perfect model. Plot y versus y_hat. It is a perfect line!  -->

### Better models make better predictions on new data

The most sensible way to test a model is to use the model to make predictions and compare those predictions to new data. After fitting the model using `stan_glm`, we would use `posterior_predict` to obtain simulations representing the predictive distribution for new cases. For instance, if we were to predict how someone's attitude changes toward immigration among Boston commuters based on political affiliation, we would want to go out and test our theories on new Boston commuters.

When thinking of generalization to new data, it is important to consider what is relevant new data in the context of the modeling problem. Some models are used to predict the future and, in those cases, we can wait and eventually observe the future and check how good our model is for making predictions. Often models are used to obtain insight to some phenomenon without immediate plan for predictions. This is the case with our Boston commuters example. In such cases, we are also interested whether learned insights from on part of the data generalizes to other parts of the data. For example, if we  know how political attitudes informed future immigration stances in *Boston commuters*, we may want to know if those same conclusions could generalize to train commuters in different locations. 

Even if we had detected clear problems with our predictions, this would not necessarily mean that there is anything wrong with the model as fit to the original dataset. However, we would need to understand it further before generalizing to other commuters. 

Often, we would like to evaluate and compare models without waiting for new data. One can simply evaluate predictions on the observed data. But since these data have already been used to fit the model parameters, these predictions are optimistic.

In cross validation, part of the data is used to fit the model and the rest of the data—the hold-out set—is used as a proxy for future data. When there is no natural prediction task for future data, we can think of cross validation as a way to assess generalization from one part of the data to another part.

In any form of cross validation, the model is re-fit leaving out one part of the data and then the prediction for the held-out part is evaluated. In the next section, we will look at a type of cross validation called leave-one-out (LOO) cross validation.


## Comparing models in practice

To compare models without waiting for new data, we evaluate predictions on the observed data. However, due to the fact that the data has been used to fit the model parameters, our predictions are often optimistic when assessing generalization. 

In cross validation, part of the data is used to fit the model, while the rest of the data is used as a proxy for future data. We can think of cross validation as a way to assess generalization from one part of the data to another part. How do we do this? 

<!-- In leave-one-out cross validation, we want to perform inferences excluding each data point i, one at a time. In the above expression, excluding that one data point is equivalent to multiplying the posterior distribution by the factor 1/p(yi |θ). The LOO posterior excluding point i is written as p(θ|y−i) = p(θ|y)/p(yi|θ),andtheLOOdistributioniscomputedbytakingtheposteriorsimulations for θ obtained from stan_glm and giving each simulation a weight of 1/p(yi |θ). This set of weighted simulations is used to approximate the predictive distribution of yi , the held-out data point. -->

We can hold out individual observations, called leave-one-out (LOO) cross validation; or groups of observations, called leave-one-group-out cross validation; or use past data to predict future observations, called leave-future-out cross validation. When we perform cross validation, the model is re-fit leaving out one part of the data and then the prediction for the held-out part is evaluated. 

For our purposes, we will be performing cross validation using leave-one-out (LOO) cross validation. 

### Cross validation using `loo()`

<!-- This is a rough draft. Everything needs to be expanded, but the basics are here. --> 

To compare models using leave-one-out (LOO) cross validation, one piece of data is excluded from our model. The model is then re-fit and makes a prediction for the missing piece of data. The difference between the predicted value and the real value is calculated. This process is repeated for every row of data in the dataset. 

In essence: *One piece of data is excluded from our model, the model is re-fit, the model attempts to predict the value of the missing piece, we compare the true value to the predicted value, and we assess the accuracy of our model's prediction.* This process occurs for each piece of data, allowing us to assess the model's accuracy in making predictions.

To perform leave-one-out(LOO) cross validation, we will be using the function `loo()` from an R package. This is how we will determine which model is superior for our purposes. 

First, we will refamiliarize ourselves with our first model, `fit_liberal`. 

```{r}
fit_liberal
```

Now, we will perform `loo()` on our model and look at the results. 

```{r}
loo_liberal <- fit_liberal %>% 
  loo()

loo_liberal
```

What does any of this mean? 

- `elpd_loo` is the estimated log score along with a standard error representing uncertainty due to using only 115 data points.
- `p_loo` is the estimated “effective number of parameters” in the model. 
- `looic` is the LOO information criterion, −2 elpd_loo, which we compute for comparability to deviance.

For our purposes, we mostly need to focus on `elpd_loo`. Let's explain, in more depth, what this information means. 

Basically, when we run `loo()`, we are telling R to take a piece of data out of our dataset, re-estimate all parameters, and then predict the value for the missing piece of data. The value for `elpd_loo()` is based off of how *close* our estimate was to the truth. Therefore, `elpd_loo()` values inform us of the effectiveness of our model in predicting data it has not seen before. The ideal value for `elpd_loo()` is 0, meaning that our predictions are perfect. This is, of course, not going to be the case. We say this to point out that, the closer our value of `elpd_loo` is to 0, the more effective our model performs.

Conclusion: **the higher our value for `elpd_loo`, the better our model performs.** This means that, when comparing models, we want to select the model with the higher value for `elpd_loo`. Hint: the higher value will be the the value closest to 0, as the results for `elpd_loo()` are negative. 

Let's turn our attention to our second model. To begin, let's observe the qualities of `fit_att_start` once again. 

```{r}
fit_att_start
```

Great! Now, let's perform `loo()` on this model. 

```{r}
loo_att_start <- fit_att_start %>% 
  loo()

loo_att_start
```

The `elpd_loo` value for this model is -201.7. This is higher than the `elpd_loo` for `att_liberal`, implying that **this** model is superior. However, we can't see our estimates together. Is there a simpler way to calculate which model is better? 

Actually, yes! Using the function `loo_compare()`, we can compare the models directly. 

### Comparing models using `loo_compare()`

To compare the two models directly, we can use the function `loo_compare` with our two `loo` objects created above. This will calculate the difference in `elpd_loo()` between our models for us, making our job easier:

```{r}
loo_compare(loo_att_start, loo_liberal)
```


The value for `elpd_diff` is equal to the difference in `elpd_loo` between our two models: elpd_loo for loo_att_start - elpd_loo for loo_liberal = -279.4 - (-201.7) = -77.7. The `se_diff` shows that the difference has a standard error of 12.0. This shows that, even when accounting for standard error, the better model is clear: `fit_att_start`. Therefore, the attitude at the start of the `trains` study is more significant to predicting final attitude when compared with the variable `liberal`, which is an analog for political affiliation. 

As we have seen,`loo_compare` is a shortcut for comparing two models. When you are deciding between two models, `loo_compare()` is a great way to simplify your decision. 

What do we do when the value of `loo_compare()` is small? As a general practice, differences smaller than four are hard to distinguish from noise. In other words: when `elpd_diff` is less than 4, there is no advantage to one model over the other. 

<!-- ### Selecting the rows -->

## Testing is nonsense 

As always, it is important to look at the practices of other professionals and the reasons we may choose not to follow those tactics. For instance, our continued problem with hypothesis testing. In hypothesis testing, we assert a null hypothesis $H_0$ about our data and an alternative hypothesis $H_a$. 

When performing hypothesis testing, we either reject the hypothesis or we do not reject it. The qualifications for rejecting are met if the 95% confidence interval excludes the null hypothesis. If the hypothesis *is* included in our 95% confidence interval, we do not reject it. In the case of “insignificant” results, with p > 0.5, we also can’t "reject" the null hypothesis. However, this does not mean that we accept it.

The premise of hypothesis testing is to answer a specific question -- one that may not even be particularly relevant to our understanding of the world -- about our data. So, what are our problems with hypothesis testing? 
- Rejecting or not rejecting hypotheses doesn't helps us to answer real questions.
- The fact that a difference is not “significant” has no relevance to how we use the posterior to make decisions.
- Statistical significance is not equal to *practical* importance. 
- There is no reason to **test** when you can summarize by providing the full posterior probability distribution.


## Parallel lines

<!-- att_end ~ liberal + att_start. Same thing as chapters before. Show graphic of parallel lines. Cardinal virtues as always. And, extra, at the end we can use loo_compare to see if this model is better than simpler model of att_end ~ att_start. -->


### Comparing complex and simple models 

Complex: 
```{r}
fit_first <- stan_glm(formula = att_end ~ liberal + att_start,
                  data = trains,
                  refresh = 0)
```

Simple: 
```{r}
fit_simple<- stan_glm(formula = att_end ~ att_start,
                  data = trains,
                  refresh = 0)
```

Make loo() regression for both:
```{r}
loo_first <- fit_first %>% 
  loo()

loo_simple <- fit_simple %>% 
  loo()
```

Perform loo_compare:
```{r}
loo_compare(loo_first, loo_simple)
```

## Summary






