# Python {-}


```{r, include=FALSE}
library(reticulate)
```

<!-- DK: Only evaluate code chunks which you are sure will work for everyone. We will discuss this command use_condaenv("C:/MINICO~1/envs/r-reticulate") Monday. -->




## Getting Started

This additional Python section is primarily for those who have done the previous chapters and associated tutorials and want to learn how to do what they did in previous chapters in Python. Before moving forward, you have to install the *reticulate* library in R. This library includes a comprehensive set of tools for interoperability between R and Python. After installing the *reticulate* library, running the following code will allow you to run Python codes in R.

```{r, eval=FALSE}
library(reticulate)
```

There are variety of ways to write python codes in Rstudio, such as creating code chunks in R markdown or creating an interactive Python console within R. For further details about *reticulate*, see [this documentation](https://rstudio.github.io/reticulate/)

### Installing Python 3

Throughout this chapter, you will use Python3 and Rstudio. Assuming you have already did the previous chapters and installed Rstudio, all you need to do is download Python3. Python is a programming language that runs computations and is highly functional for data science.

Initial step is to Download and install Python 3.

1.**installing Python 3**: [Download and install Python 3.](https://www.python.org/downloads/)

### Package Installation 

Assuming you know the how the package system works in R, package system works very similar in Python. 
You need to:

  1. *Install the package*: 
     This is like installing an app on your phone. Most packages are not installed by default when you install Python. Thus, if you want to use a package for the first time, you need to install it. Once you’ve installed a package, you likely will not install it again unless you want to update it to a newer version.
 
 2. *“Load” the package*: 
    “Loading” a package is like opening an app on your phone. Packages are not “loaded” by default when you start Rstudio. You need to “load” each package you want to use every time you restart Rstudio.


One way to install the package is through reticulate library. For example, to import numpy- a popular library used to statistically analyse a data set- write the following code in the console. 

```{python, eval=FALSE}
py_install(numpy)
```

After writing this step, hit enter on your keyboard. You should also install two more packages we will use in this chapter: “Matplotlib’ and ‘Pandas.’

### Package loading

To use the installed packages, we have to load them, and we do this by using the import statement. If you import a package in one cell, you do not need to import it again in another cell as long as you do not restart the Jupyter Notebook.
For example, to load the numpy package, run the following code in a code cell in the notebook.  

```{python, eval=FALSE}
import numpy as np
```

To save time, when we want to load a library, we usually import it under a different name. Hence, ‘as np’ is generally called when importing numpy. 

### Package Use

You have to load each package you want to use every time you start Rstudio. If you don’t load a package before attempting to use one of its features, you may see an error message like:


```{python,error=TRUE}
simple_values = [1,2,3]
print(numpy.mean(simple_values))
```

With this Error, Python is telling you that you are trying to use a function that has not been loaded yet. If you load the package and then run the code, it will work without an error, as shown below. 

```{python}
import numpy as np
simple_values = [1,2,3]
print(np.mean(simple_values))
```

### Summary

You should have done the followings: 
  * Installed the latest versions of Python 3.
  * Installed the *reticulate* package in R. 
  * Installed the following three packages: 
    1. Numpy
    2. Matplotlib
    3. Pandas






