---
title: "Probability Questions"
author: "Vivian Zhang"
date: "6/12/2020"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

STAT 110 interesting questions:

- You're ordering two pizzas. A pizza can be small, medium, large, or extra large, with any combination of 8 possible toppings (getting no toppings is allowed, as is getting all 8). How many possibilities are there for your two pizzas?

- According to the CDC, men who smoke are 23 times more likely to develop lung cancer than men who don't smoke. Also according to the CDC, 21.6% of men in the U.S. smoke. What is the probability that a man in the U.S. is a smoker, given that he develops lung cancer?

- Researcher wants to estimate the percentage of people who have used illegal drugs by surveying them! But people lie when asked sensitive questions about illegal drugs. Thus, the researcher uses a randomized response. They fill a hat with slips of paper, each of which says either "I have used illegal drugs" or "I have not used illegal drugs." Let p be the proportion of slips of paper that say "I have used illegal drugs" (p chosen by the researcher in advance).

Each participant chooses a random slip from the hat and answers yes or no truthfully. The researcher doesn't know what slip the person had. Let y be the probability that a participant will say "yes" and d be the probability that a participant has used drugs. Find y, in terms of d and p. What's the worst possible p that the researcher could make?

Suppose that proportion p of the slips say "I have used illegal drugs" but the remaining 1-p say "I was born in the winter". Assume 1/4 people are born in winter and season of birth is independent of whether a person used illegal drugs. Find d, in terms of y and p.

<!-- e) Need to make some/all of these question such that they can not be solved analytically. Or, at least, have a second part of the question like that. -->


<!-- e) In our main examples, we treat n as fixed and show the joint distribution of p and the number of heads H. All good. But we could also treat H as fixed and show the joint distribution between n and p. Two ways to do that. First, hard programming way is to flip the count until you get H heads. Second, easy simulation way, flip H * 1000 times for each replication and then, from that tibble, throw away all the rows which don't have exactly H heads. Either way leaves you with a tibble which is the joint pdf. -->

<!-- f) Another problem set question would take calculate the posterior for ranges, not just for point estimates. What is the marginal distribution for p if we see, say, an even number of heads, rather than just for 8 heads? Easier, perhaps, is to think of the compact interval equivalent. What is p if we see at least 8 heads? Or, if we calculate 1,001 ps, what does the data look like if 0.52 < p < 0.55?  -->

<!-- g) Problem set question might be: You have two coins, A and B. You know that p, the probability of heads, can only be 0, 0.1, 0.2, ..., 1 --- nothing in between. You flip both counts 20 times. A shows 8 heads, B shows 7. What are the odds that B will have at least as many heads as A if you flip the coins another 30 times? Extra credit: Same question, but you may assume that the value of p for A and B is different. Also, show a cool histogram of the posterior results, perhaps based on an example we present. -->

<!-- f) The pancake problem from chapter 14 from rethinking is interesting, perhaps as a class exercise. There are three pancakes: one burned on both sides (BB); one unburnt on both sides (UU); one burned on one side and unburnt on the other (BU). If I serve you a pancake which is burnt on the top (which is the only side you can see), what are the odds that it is burnt on the other side? Intuition says: 1/2. You don't have UU pancake, so it is 50/50 which other pancake you have. But the truth is 2/3s, because you need to think in terms of sides of a pancake, not the whole pancake. There are 3 sides which could be hidden (the UU is still impossible), but 2 of them are B, so the odds of 2/3s. Intuition does not work well here, but simulation does. (I think? Does the Bayes Scatterplot have three rows and two columns?)  -->

```{r Question 1 of the 2011 Midterm}

# A new treatment for a disease is being tested, to see whether it is better
# than the standard treatment. The existing treatment is e↵ective on 50%
# of patients. It is believed initially that there is a 2/3 chance that the new
# treatment is effective on 60% of patients, and a 1/3 chance that the new
# treatment is effective on 50% of patients. In a pilot study, the new treatment
# is given to 20 random patients, and is e↵ective for 15 of them.
# Given this information, what is the probability that the new treatment
# is better than the standard treatment (as an unsimplified number)?

sims <- 10000000

# 60%

asims <- ceiling(2/3 * sims)

# 50%

bsims <- floor(1/3 * sims)

# Assignment 

atest <- matrix(rbinom(n = 20 * asims, size = 1, prob = 0.6), nrow = asims, ncol = 20)

btest <- matrix(rbinom(n = 20 * bsims, size = 1, prob = 0.5), nrow = bsims, ncol = 20)

# Analysis 

aresult <- data.frame(successes = rowSums(atest))

bresult <- data.frame(successes = rowSums(btest))

anum <- aresult %>% 
  filter(successes == 15) %>% 
  summarize(total = n())

bnum <- bresult %>% 
  filter(successes == 15) %>% 
  summarize(total = n())

print(anum / (anum + bnum))

```

```{r widgets}

# A widget inspector inspects 12 widgets and finds that exactly 3 are defective.
# Unfortunately, the  widgets  then  get  all  mixed  up  and  the  inspector
# has to find the 3 defective widgets again by testing widgets one by one.(a) 
# Find the probability that the inspector will now have to test at least 9 
# widgets.(b) Find the probability that the inspector will now have to test at 
# least 10 widgets.

sims <- 100

widgets <- tibble(
  quality = c(
    rep(0, 3), # 0 for defective
    rep(1, 9) # 1 for good
))

# Needs to be fixed

mysample <- map(widgets, ~rep_sample_n(widgets, sims, replace=TRUE))

test <- matrix(mysample$quality, nrow = sims, ncol = 12)

get_widgets <- function(n) {
  sample(widgets, n, replace=TRUE)
}

get_widgets(12)

x <- tibble(results = map(1:100, ~ roll_alpha_dice(sims)))

roll_alpha_dice <- function(N){
  map_chr(1:sim, ~ get_letters(12))
}


x <- tibble(results = map(1:100, ~ roll_alpha_dice(sims)))

mydata <- map_lgl(widgets, ~rep_sample_n(widgets, size = 12,
                                         replace = FALSE, reps = sims))

mydata <- map_df(sizes, ~ rep_sample_n(all_students, size =., replace = FALSE, reps = 5000), .id = "size")

test <- matrix(rbinom(n = 12 * sims, size = 1, prob = 1/4), 
               nrow = sims, ncol = 12)

# 1 means defective, 0 means good

aresult <- data.frame(successes = rowSums(test))



?rbinom

```
